\documentclass{article}
\usepackage{fullpage}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{times}
\usepackage{graphicx}
\usepackage{alltt}
\usepackage{algpseudocode}

\begin{document}

%\thispagestyle{empty}

\noindent
\fbox{
\parbox{\textwidth}{
\begin{Large}
{\bf CS 577: Introduction to Algorithms\hfill Greedy Review Problems}
\end{Large}
}}

\subsection*{Problems}

\begin{enumerate}
%
% Problem 1
% taken from 1.pdf, MST
	\item 
		Let $G = (V,E)$ be a connected undirected graph.  We will call a subgraph
$(V,E')$ \emph {special} if it is connected and has exactly one cycle.

	\begin{enumerate}
		\item Show that a subgraph is special iff it can be obtained by taking a
spanning tree and adding one edge.
		\item Let $G$ have non-negative weights on its edges.  Suppose all edge
weights are distinct. Describe and analyze efficient algorithm that finds a
minimum-weight special subgraph of $G$, assuming there is one.  
	\end{enumerate}
%
% Problem 2
% taken from 6.pdf, MST in intervals

	\item Suppose we have a connected graph $G = (V, E)$. Each edge $e$ has a
time-varying edge cost given by a function $f_e(t) = a_et + b_e$ where $f_e(t)$
gives the cost of the edge at time $t$.  Observe that the set of edges
constituting the minimum spanning tree of $G$ may change over time, and that
the cost of the minimum spanning tree of $G$ can be expressed as a function of
the time $t$. Denote this function $cost_G(t)$.

Give an algorithm that takes as input: -- the graph $G$; -- the values $\{(a_e,
b_e) : e \in E\}$; -- an interval $[\ell , r]$; and returns a value of the time
$t\in [\ell,r]$ at which the minimum spanning tree has minimum cost, $\min_{t
\in [\ell,r]} cost_G(t)$. Your algorithm should run in time polynomial in the
number of nodes and edges of the graph $G$, where you may assume that
arithmetic operations on numbers can be done in constant time.

%
%Problem 3
%taken from 6.pdf
		\item Given a connected graph $G=(V,E)$ with weights $w_e$ on edges and
a spanning tree $T$, define the width of $T$ to be the weight of the maximum
weight edge in $T$.

  \begin{enumerate}
	  \item Given a target width $W$, develop a linear time algorithm to
determine whether $G$ contains a spanning tree of width at most $W$.

		 \item Develop a linear time algorithm to find the smallest $W$ such
that there exists a spanning tree $T$ with width $W$ in $G$. You may assume
that all edge weights are distinct.

    (Hint: Use your algorithm from part (a) as a subroutine, and do a
    sort of ``binary search'' over $W$.)
   
\end{enumerate}

%
%Problem 4
%Taken from 3.pdf
	\item  Consider a path between two vertices $s$ and $t$ in an undirected
weighted graph $G$. The \emph{bottleneck length} of this path is the maximum
weight of any edge in the path. The \emph{bottleneck distance} between $s$ and
$t$ is the minimum bottleneck length of any path from $s$ to $t$. (If there are
no paths from $s$ to $t$, the bottleneck distance between $s$ and $t$ is
$\infty$.)

Describe and analyze an efficient algorithm to compute the bottleneck distance between every pair of vertices
in an arbitrary undirected weighted graph. Assume that no two edges have the same
weight.


\end{enumerate}
\end{document}
