%
\documentclass{article}
\usepackage{fullpage}

\usepackage{epsfig}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amscd}
\usepackage{amsmath}
\usepackage{theorem}
\usepackage{times}
\usepackage{graphicx}
\usepackage{alltt}
\usepackage{enumerate}

\begin{document}

%\thispagestyle{empty}

\noindent
\fbox{
\parbox{\textwidth}{
\begin{Large}
{\bf CS 577: Introduction to Algorithms\hfill Divide and Conquer Review}
\end{Large}\\

\begin{large}
{\bf  \hfill }
\end{large}
}}

\subsection*{Problems}

\begin{enumerate}
\item
	For this problem, a subtree of a binary tree means any connected subgraph. A binary tree is complete if every internal node has two children, and every leaf has exactly the same depth. Describe and analyze a recursive algorithm to compute the $largest$ $complete$ $subtree$ of a given binary tree. Your algorithm should return the root and depth of the subtree.

\item
    Given an array $A[1\ldots n]$ of integers, develop an algorithm to count the number of significant inversions where given $i$ and $j$ as indexes in the array, it satisfies $i<j$ and $A[i]>2*A[j]$. A sorted array would have an inversion count of 0.
    
\item  
    Consider a $n$ by $n$ matrix of integers where each row is sorted in increasing order from top to bottom, and each column is sorted in increasing order from left to right. Design and analysis a divide and conquer algorithm similar to binary search to determine whether a given integer lies in the matrix, and if so return its row and column number assuming both row and column start with index 0 and end at index n-1. Assume that all integers in matrix are distinct

\item 
    In this problem, you are given an array $A[0\cdots n-1]$ of $n$
distinct numbers sorted in increasing order and then right-shifted
circularly by some unknown number of places. Your goal is to figure
out the amount of shift, equivalently, the position of the smallest
element in the array. (Note: we are indexing the array from $0$ to
$n-1$ here.) For example, if you are given the array
$[5,7,9,10,1,3,4]$, your answer should be $4$, which is the index of the
smallest element $1$ and is also the number of places the sorted array
$[1,3,4,5,7,9,10]$ would need to be right-shifted to produce the given array.
Develop a divide and conquer algorithm for this problem.
Hint: Try writing a procedure to find the minimum element of the list.


\end{enumerate}
\end{document}
